### 2014-11-03 splash-bp.tex
Splash BP results.

The asynchronous approach proves to be a nice improvement over the synchronous
version. Still, it is possible to do even better. Gonzalez et
al~\cite{Gonzalez+al:aistats09paraml} developed an optimal algorithm to compute
this algorithm by first building a tree and then updating the beliefs of each
node twice, first from the leaves to the root and then from the root to the
leaves. The root of this tree is the node with the highest priority (based on
belief) while the other nodes in the tree must have a non-zero priority.
Note that the priorities are updated whenever a neighbor updates their belief.
These \emph{splash trees} are built iteratively until we reach convergence.

\begin{figure}[h!]
   \begin{center}
      \includegraphics[width=6.5cm]{figures/splash_bp}
   \end{center}
   \caption{Single Source Shortest Path program code.}
   \label{splash_bp}
\end{figure}

@@

When we have several threads, every thread will generate their own trees by
taking into account the highest priority node in their own queues. This is the
reason we use \texttt{set-static} in line 1 to force all nodes to remain in the
original threads.

@@

// TREE BUILDING
// continue tree
waiting(A), token(A, All, Next) -o token(A, All, Next).
// start tree
waiting(A), *underline{@priority(A, A, P)}, P > 0.0
   -o token(A, [A], [A]).
// end tree building
token(A, All, Next), length(All) > maxnodes
   -o first-phase(A, All, reverse(All)).
// expand tree
token(A, All, [A | Next])
   -o [collect => L | Side | !edge(A, L, Side),
         0 = count(All, L),
         0 = count(Next, L),
         *underline{priority(A, L, P)}, P > 0.0,
         *underline{cpu-id(A, L, Id1)},
         *underline{cpu-id(A, A, Id2)}, Id1 = Id2 |
         send-token(A, All, Next ++ L)].

@@

// SECOND PHASE
second-phase(A, [], _)
   -o *underline{set-priority(A, 0.0)}, waiting(A), update(A).
second-phase(A, [A], Back)
   -o update(A), waiting(Back),
      waiting(A), *underline{set-priority(A, 0.0)}.
second-phase(A, [A, B | Next], Back)
   -o update(A), waiting(Back), *underline{schedule-next(B)},
      second-phase(B, [B | Next], A).
\end{Verbatim}
  \caption{Coordination code for the Splash Belief Propagation program.}
  \label{code:sbp}
\end{figure}
\normalsize

@@

>>>>

The asynchronous approach proves to be a nice improvement over the synchronous
version. Still, it is possible to do even better. Gonzalez et
al~\cite{Gonzalez+al:aistats09paraml} developed an optimal algorithm to compute
this algorithm by first building a tree and then updating the beliefs of each
node twice, first from the leaves to the root and then from the root to the
leaves. The root of this tree is the node with the highest priority (based on
belief) while the other nodes in the tree must have a non-zero priority.
Note that the priorities are updated whenever a neighbor updates their belief.
These \emph{splash trees} are built iteratively until we reach convergence.
In Fig.~\ref{splash_bp} we represent two threads creating two different splash
trees. It is encouraged to create several splash trees concurrently as long as
we have threads to create them.

\begin{figure}[h!]
   \begin{center}
      \includegraphics[width=6.5cm]{figures/splash_bp}
   \end{center}
   \caption{Creating splash trees for belief propagation. Each threads picks the
   highest priority node and creates a tree from that node. The belief values
   are updated in two phases: first from the leaves to the root and then from
   the root to the leaves.}
   \label{splash_bp}
\end{figure}

@@

Note that the \texttt{set-static} action fact is used in line 1 in order to
force all nodes to remain in the original threads, so that each thread has a
well defined area of nodes to build splash trees on.

@@

// TREE BUILDING
// continue tree
waiting(A), token(A, All, Next) -o token(A, All, Next).
// start tree
waiting(A), *underline{@priority(A, A, P)}, P > 0.0
   -o token(A, [A], [A]).
// end tree building
token(A, All, Next), length(All) > maxnodes
   -o first-phase(A, All, reverse(All)).
// expand tree
token(A, All, [A | Next])
   -o [collect => L | Side | !edge(A, L, Side),
         0 = count(All, L),
         0 = count(Next, L),
         *underline{@priority(A, L, P)}, P > 0.0,
         *underline{@cpu-id(A, L, Id1)},
         *underline{@cpu-id(A, A, Id2)}, Id1 = Id2 |
         send-token(A, All, Next ++ L)].

@@

// SECOND PHASE
second-phase(A, [], _)
   -o *underline{set-priority(A, 0.0)}, waiting(A), update(A).
second-phase(A, [A], Back)
   -o update(A), waiting(Back),
      waiting(A), *underline{set-priority(A, 0.0)}.
second-phase(A, [A, B | Next], Back)
   -o update(A), waiting(Back), *underline{schedule-next(B)},
      second-phase(B, [B | Next], A).
\end{Verbatim}
  \caption{Coordination code for the Splash Belief Propagation program. When a
     linear fact is prefixed by \texttt{@}, it indicates that the fact is going
  to be re-derived.}
  \label{code:sbp}
\end{figure}
\normalsize

@@

\begin{figure*}[]
   \begin{center}
      \subfloat[]{\includegraphics[width=5cm]{results/system_belief-propagation-400.png}}
      \subfloat[]{\includegraphics[width=5cm]{results/system_splash-bp-400.png}}
      \subfloat[]{\includegraphics[width=5cm]{results/system_improve_belief-propagation-400.png}}
   \end{center}
   \caption{Memory usage and scalability of the regular and coordinated versions
      of MiniMax.}
   \label{results:splash_bp}
\end{figure*}
